---
title: "Inspect data of DrugCell"
author: "Chunguang Lai"
date: "2022-02-18"
output:
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../")

```
# 1.Preparation 
## 1.1 Enviroment setup
```{r, error = TRUE}
requiredPackages = c('here','ggplot2','compare')
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}
```


## 1.2 Load drug infomation

```{r, error = TRUE}

compoundname=read.csv2(here("data","compound_names.txt"),header=TRUE,sep="\t")
# CN1C=NC2=C1C=C(C(=C2F)NC3=C(C=C(C=C3)Br)Cl)C(=O)NOCCO
drugcell_all=read.csv2(here("data","drugcell_all.txt"),header=FALSE,sep="\t")

# table(grepl("CN1C=NC2=C1C=C(C(=C2F)NC3=C(C=C(C=C3)Br)Cl)C(=O)NOCCO",drugcell_all[,2],fixed=TRUE))
```

# 2. Plot distribution of mutations per cell line as Fig S1E
## plot Fig S1E, Histogram of the number of mutations per cell line used for model training, matched figure S1E
```{r, error=TRUE}

cellname=read.csv2(here("data","cell2ind.txt"),header=FALSE,sep="\t")
cellmutation=read.csv2(here("data","cell2mutation.txt"),header=FALSE,sep=",")
genename=read.csv2(here("data","gene2ind.txt"),header=FALSE,sep="\t")
distribution_cellmutation=as.data.frame(apply(cellmutation,1,FUN=sum))
distribution_drugmutation=as.data.frame(apply(cellmutation,2,FUN=sum))


distribution_cellmutation1=as.data.frame(t(rbind(c(1:length(distribution_cellmutation[,1])),(distribution_cellmutation[,1]))))

ggplot(data.frame(distribution_cellmutation1), aes(x=V2))+xlab("# mutation")+ylab("# cell lines")+ scale_x_continuous(limits=c(1,1335),trans='log10')+
  geom_histogram(color="darkblue", fill="lightblue",binwidth=0.1
  )


```

# 3. Check number of cell lines tested with specific drugs

## 3.1 Paclitaxel: total 997 pairs for Paclitaxel, 2 different structures, which have 239, 758 cell lines respectively, 239 is same as Fig 3F's number
```{r, error = TRUE}
compoundname[which(grepl("Paclitaxel",compoundname[,3],ignore.case = TRUE)),]
# first pacilitaxel row 265, has 239 pairs/cell lines which match the number of Figure 3F
drugcell_pacilitaxel1=drugcell_all[drugcell_all[,2]=="CC1=C2[C@H](C(=O)[C@@]3([C@H](C[C@@H]4[C@]([C@H]3[C@@H]([C@@](C2(C)C)(C[C@@H]1OC(=O)[C@@H]([C@H](C5=CC=CC=C5)NC(=O)C6=CC=CC=C6)O)O)OC(=O)C7=CC=CC=C7)(CO4)OC(=O)C)O)C)OC(=O)C",]
length(table(drugcell_pacilitaxel1[,1]))

# second pacilitaxel row 359, has 758 pairs/cell liens, I don't know what it is

drugcell_pacilitaxel2=drugcell_all[drugcell_all[,2]=="CC1=C2[C@H](C(=O)[C@@]3([C@H](C[C@@H]4[C@](C3[C@@H]([C@@](C2(C)C)(C[C@@H]1OC(=O)[C@@H]([C@H](C5=CC=CC=C5)NC(=O)C6=CC=CC=C6)O)O)OC(=O)C7=CC=CC=C7)(CO4)OC(=O)C)O)C)OC(=O)C",]
length(table(drugcell_pacilitaxel2[,1]))

```


## 3.2 selumetinib: One mono test, and several combinations for selumetinib, 769 cell lines doesn't match Figure 3B's n=382
```{r, error = TRUE}

# 8948 of combinations involved selumetinib
table(grepl("CN1C=NC2=C1C=C(C(=C2F)NC3=C(C=C(C=C3)Br)Cl)C(=O)NOCCO",drugcell_all[,2],fixed=TRUE))


compoundname[which(grepl("selumetinib",compoundname[,3],ignore.case = TRUE)),]

# Selumetinib at row 355, has 788 pairs, 769 cell lines, dosen't match Figure 3B's n=382


length(which(drugcell_all[,2]=="CN1C=NC2=C1C=C(C(=C2F)NC3=C(C=C(C=C3)Br)Cl)C(=O)NOCCO"))

drugcell_selumetinib=drugcell_all[(which(drugcell_all[,2]=="CN1C=NC2=C1C=C(C(=C2F)NC3=C(C=C(C=C3)Br)Cl)C(=O)NOCCO")),]
length(table(drugcell_selumetinib[,1]))

```

## 3.3 JQ1: one mono, and several combinations for JQ1, mono is row 661, 839 doesn't match Figure 3D's n=860
```{r, error = TRUE}

compoundname[which(grepl("JQ1",compoundname[,3],ignore.case = TRUE)),]

# JQ1 row 661, has 849 pairs, 808 cell lines, dosen't match Figure 3D's n=860

length(which(drugcell_all[,2]=="CC1=C(SC2=C1C(=N[C@H](C3=NN=C(N32)C)CC(=O)OC(C)(C)C)C4=CC=C(C=C4)Cl)C"))
drugcell_JQ1=drugcell_all[(which(drugcell_all[,2]=="CC1=C(SC2=C1C(=N[C@H](C3=NN=C(N32)C)CC(=O)OC(C)(C)C)C4=CC=C(C=C4)Cl)C")),]
length(table(drugcell_JQ1[,1]))


# 5279 pairs, 839 cell lines involved JQ1 in total, still doesn't match 860
JQ1="CC1=C\\(SC2=C1C\\(=N\\[C@H\\]\\(C3=NN=C\\(N32\\)C\\)CC\\(=O\\)OC\\(C\\)\\(C\\)C\\)C4=CC=C\\(C=C4\\)Cl\\)C"
table(grepl(JQ1,drugcell_all[,2]))
drugcell_JQ1_all=drugcell_all[(grepl(JQ1,drugcell_all[,2])),]
length(table(drugcell_JQ1_all[,1]))

```




# 4. Check number of cell lines with specific gene mutation

## 4.1 BRAF mutations, but 175 doesn't match with figure 3A's n=229

```{r, error = TRUE}
# BRAF is row 288
genename[grepl("BRAF",genename[,2])==TRUE,]

# but it only has 175 mutated cell lines, not 229
table(data.frame(cellmutation[,288]))

```



## 4.2 EGFR,BRAF,LKB1 mutations, 311 in total, still doesn't match with Fig 3C's 460

```{r, error = TRUE}
# 802
genename[grepl("EGFR",genename[,2])==TRUE,]
# 1513
genename[grepl("LKB1",genename[,2])==TRUE,]

# table(data.frame(cellmutation[,802]))
# table(data.frame(cellmutation[,1513]))
# table(data.frame(cellmutation[,288]))

#  EGFR,BRAF,LKB1 mutations have 311 cell lines in total, doesn't match 460
table(rowSums(data.frame(cellmutation[,c(288,802,1513)])))
1225-table(rowSums(data.frame(cellmutation[,c(288,802,1513)])))[[1]]


```

# 5. Inspect cell2mutation_fixed.txt, and find that this dataset has 1224 cell ines, which is lack first cell line compared with 1225 of cell2mutation.txt. But still get 311 as last section

```{r, error=TRUE}
cellmutation2=read.csv2(here("data","cell2mutation_fixed.txt"),header=FALSE,sep=",")


compare_0.001=function(x){return(x>0.001)}
cellmutation3=data.frame(sapply(cellmutation2,FUN=compare_0.001))
cellmutation3=data.frame(sapply(1*cellmutation3,FUN=as.integer))
dim(cellmutation3)

table(rowSums(data.frame(cellmutation3[,c(288,802,1513)])))
1224-table(rowSums(data.frame(cellmutation3[,c(288,802,1513)])))[1]


# prove that cell2mutation.txt only has 1st row as extra cell line.
cellmutation4=cellmutation[2:dim(cellmutation)[1],]
comparison=compare(cellmutation4,cellmutation3,allowAll=TRUE)
comparison[1]
```